Psychological Needs During Intergroup Contact: Three Experience Sampling Studies

### 2022-11-07 manuscript.tex
app str added
app-b tbl & fgr
app-a ttl
app A written

% add packages
\usepackage[american]{babel}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{hyperref}
\usepackage[style=apa, sortcites=true, sorting=nyt, backend=biber, natbib=true, uniquename=false, uniquelist=false, useprefix=true]{biblatex}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{setspace,caption}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{lipsum}
\usepackage{soul}
\usepackage{xcolor}
\usepackage{fourier}
\usepackage{stackengine}
\usepackage{scalerel}
\usepackage{fontawesome5}
\usepackage[normalem]{ulem}
% \usepackage{longtable}
\usepackage{amsmath}
\usepackage{ntheorem}
\usepackage{afterpage}
\usepackage{float}
\usepackage{array}
\usepackage{censor}
\usepackage{pdflscape}
\usepackage{lscape}
\usepackage{pdfpages}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{makecell}
\usepackage{tabu}
% Path Diagrams
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
\tikzset{mynode/.style={draw,text width=1in,align=center} }
\usetikzlibrary{positioning}

@@

@@

\section{The Present Research}
The aim of this paper is essentially threefold. First, our study is among the first to test the fundamental tenets of intergroup contact and Allport's conditions in real-life extensive longitudinal data. Second, we test whether the fulfillment of situational psychological needs is meaningfully related to more positive outgroup attitudes following intergroup interactions. And third, we test the role of interaction quality in  the relationship of psychological need fulfillment and outgroup attitudes. Based on this we formulated three main hypotheses:
\begin{enumerate}[label=H\arabic*:]
    \item Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes across extensive longitudinal data.
    \item Based on Allport’s optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data.
    \item Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data. We also expect situational need fulfillments to work at least as well as Allport's optimal contact conditions in predicting outgroup attitudes.
\end{enumerate}
Beyond these main hypotheses, we have formulated several sub-hypotheses that form the basis for our analysis plan (these sub-hypotheses also include several robustness checks and are available in Appendix \ref{app:AppendixHypotheses}).

@@

Given the more complex data structure of the extensive longitudinal data, we opted to test most of our hypotheses using multilevel regression models, where the measurement occasions (level 1) were nested within the participants (level 2). Given that more complex dynamic modeling procedures are not necessary to test these more basic hypotheses, we will focus on the contemporaneous effect of the data (e.g., need fulfillment during the intergroup interaction predicts outgroup attitudes following the interaction). Such an approach is also tolerant to missing data and uneven case numbers within participants. As is common with multilevel analyses, we use a hierarchical modeling approach and always report the final model in-text (for the full modeling process see Supplementary Material B).

@@

% Contact hypothesis in ESM data [inconsistent in aggregate but consistent in ML analysis]
When considering the results of the three studies jointly, we found mixed results for the basic intergroup contact hypothesis. More specifically, we observe that separating between-participant and within-participant effects is crucial. For between-subjects effects (using aggregated reports), the positive effect of more outgroup contact did not depend on the average interaction quality. These aggregate findings are surprising because most past cross-sectional studies have also considered an aggregate measurement of interaction quality (albeit not one from actual daily reports). The absence of this aggregate effect could underline the fact that cross-sectional retrospective data might be misleading because (1) it presents a mixture of within and between-subjects effects \citep[][]{Hamaker2020}, or (2) suffers from recall biases (e.g., where times with no outgroup interactions are undervalued by participants during the retrospective evaluations).\footnote{It should be noted that such inconsistencies with past research might in part be a data artefact (e.g., because most people reported substantially more measurements during which they did not have an outgroup interaction). However, the models we employed should in principle be resistant to such imbalances.} In support of this possibility, once we used multilevel modeling to separate effects into a between-subjects and within-subjects part, we found that when having an interaction with an outgroup member (vs. not having an outgroup interaction) is a strong predictor of outgroup attitudes across all three studies (even when taking other non-outgroup interactions into account; also see Figure \ref{fig:ContactHypothesis}B). We thus find that the main benefits of intergroup contact remain visible across the three studies. However, the intensive longitudinal data of natural intergroup contacts also offers a novel opportunity to explore how people recall and cognitively aggregate outgroup interactions. Future studies may want to ask participants explicitly to evaluate the number and quality of outgroup interactions following a phase of intensive longitudinal data collection to disentangle the aggregation differences.

@@

\input{Tables/mdlContactGeneralLong}
\input{Tables/mdlTheorylong}
\input{Tables/mdlRobustlong}

% Figures
\begin{figure}
  \caption{Path Diagrams Situational Needs Model across Studies}
  \label{fig:MainPaths}
  \begin{center}
    \input{PathModels.tex}
  \end{center}
  \caption*{Note: \\
  (A) Study 1; (B) Study 2; (C) Study 3.\\
  General: Path coefficients are standardized regression coefficients. Statistical significance markers are based on the unstandardized regression results (as presented in Table \ref{tab:intergroupNeedsTblLong}).\\
  **** p < .0001, *** p < .001, ** p < .01, * p < .05}
\end{figure}

@@

\section{Appendix A: Hypotheses}
\label{app:AppendixHypotheses}

\begin{hyp}[H\ref{hyp:contact}] \label{hyp:contact}
Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes.
\end{hyp}

\begin{subhyp}[H\ref{hyp:contactFreq}] \label{hyp:contactFreq}
\addtolength{\leftskip}{2.5em}
Participants with more intergroup interactions should have more favorable outgroup attitudes.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:contactDummy}] \label{hyp:contactDummy}
\addtolength{\leftskip}{2.5em}
Outgroup attitudes should be more positive after an intergroup interaction compared to a non-outgroup interaction.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:contactFreqQual}] \label{hyp:contactFreqQual}
\addtolength{\leftskip}{2.5em}
Participants with more intergroup interactions should have more favorable outgroup attitudes depending on the average interaction quality.
\end{subhyp}

\begin{hyp}[H\ref{hyp:AllportsConditions}] \label{hyp:AllportsConditions}
Based on Allport's optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
\end{hyp}

\setcounter{subhyp}{0}
\begin{subhyp}[H\ref{hyp:AllportsPred}] \label{hyp:AllportsPred}
\addtolength{\leftskip}{2.5em}
Based on Allport's optimal contact conditions, outgroup attitudes should be more favorable after intergroup interactions with equal status, common goals, collaboration, and structural support.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:AllportsQuality}] \label{hyp:AllportsQuality}
\addtolength{\leftskip}{2.5em}
Based on past research on the role of interaction quality, interaction quality should be more perceived as more favorable after intergroup interactions with equal status, common goals, collaboration, and structural support.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:AllportsQualityMediation}] \label{hyp:AllportsQualityMediation}
\addtolength{\leftskip}{2.5em}
Based on past research on the role of interaction quality, the variance explained in outgroup attitudes by Allport's optimal contact should to a large extent be assumed by interaction quality.
\end{subhyp}

\begin{hyp}[H\ref{hyp:keyNeed}] \label{hyp:keyNeed}
Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
\end{hyp}

\setcounter{subhyp}{0}
\begin{subhyp}[H\ref{hyp:keyNeedPred}] \label{hyp:keyNeedPred}
\addtolength{\leftskip}{2.5em}
Outgroup attitudes should be more favorable after intergroup interactions with high key-need fulfillment.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:keyNeedQual}] \label{hyp:keyNeedQual}
\addtolength{\leftskip}{2.5em}
Interaction Quality should be perceived as more positive after intergroup interactions with higher key need fulfillment.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:keyNeedMediation}] \label{hyp:keyNeedMediation}
\addtolength{\leftskip}{2.5em}
The variance explained in outgroup attitudes by key need fulfillment should to a large extend be assumed by interaction quality.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:keyNeedContactType}] \label{hyp:keyNeedContactType}
\addtolength{\leftskip}{2.5em}
The effect of key need fulfillment on outgroup attitudes should be specific to intergroup interactions and not be due to need fulfillment in general. Thus, the effect of key need fulfillment on outgroup attitudes should be stronger for intergroup interactions than for ingroup interactions. 
\end{subhyp}

\begin{subhyp}[H\ref{hyp:keyNeedSDT}] \label{hyp:keyNeedSDT}
\addtolength{\leftskip}{2.5em}
The effect of key need fulfillment on outgroup attitudes should persist even when taking other fundamental psychological needs into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for autonomy, competence, and relatedness fulfillment during the interaction (cf., self-determination theory). 
\end{subhyp}

\begin{hyp}[H\ref{hyp:comparison}] \label{hyp:comparison}
Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict outgroup attitudes at least as well as Allport's conditions.
\end{hyp}

\setcounter{subhyp}{0}
\begin{subhyp}[H\ref{hyp:compModel}] \label{hyp:compModel}
\addtolength{\leftskip}{2.5em}
The need model (H\ref{hyp:keyNeedPred}) should predict more variance in outgroup attitudes than the model based on Allport's conditions (H\ref{hyp:AllportsPred}).
\end{subhyp}

\begin{subhyp}[H\ref{hyp:compTogether}] \label{hyp:compTogether}
\addtolength{\leftskip}{2.5em}
The  effect of key need fulfillment on outgroup attitudes should  persist even when taking other Allport's conditions into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for equal status, common goals, collaboration, and structural support.  
\end{subhyp}

\section{Appendix B: Robustness}
\label{app:AppendixRobustness}

>>>>

% add packages
\usepackage[american]{babel}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{hyperref}
\usepackage[style=apa, sortcites=true, sorting=nyt, backend=biber, natbib=true, uniquename=false, uniquelist=false, useprefix=true]{biblatex}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{setspace,caption}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{lipsum}
\usepackage{soul}
\usepackage{xcolor}
\usepackage{fourier}
\usepackage{stackengine}
\usepackage{scalerel}
\usepackage{fontawesome5}
\usepackage[normalem]{ulem}
% \usepackage{longtable}
\usepackage{amsmath, nccmath}
\usepackage{mdframed}
\usepackage{ntheorem}
\usepackage{afterpage}
\usepackage{float}
\usepackage{array}
\usepackage{censor}
\usepackage{pdflscape}
\usepackage{lscape}
\usepackage{pdfpages}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{makecell}
\usepackage{tabu}
% Path Diagrams
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
\tikzset{mynode/.style={draw,text width=1in,align=center} }
\usetikzlibrary{positioning}

@@

% ref labels
\newcommand{\fgrref}[2][]{\hyperref[#2]{Figure \ref*{#2}#1}}
\newcommand{\tblref}[2][]{\hyperref[#2]{Table \ref*{#2}#1}}
\newcommand{\appref}[2][]{\hyperref[#2]{Appendix \ref*{#2}#1}}

@@

\section{The Present Research}
The aim of this paper is essentially threefold. First, our study is among the first to test the fundamental tenets of intergroup contact and Allport's conditions in real-life extensive longitudinal data. Second, we test whether the fulfillment of situational psychological needs is meaningfully related to more positive outgroup attitudes following intergroup interactions. And third, we test the role of interaction quality in  the relationship of psychological need fulfillment and outgroup attitudes. Based on this we formulated three main hypotheses:
\begin{enumerate}[label=H\arabic*:]
    \item Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes across extensive longitudinal data.
    \item Based on Allport’s optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data.
    \item Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data. We also expect situational need fulfillments to work at least as well as Allport's optimal contact conditions in predicting outgroup attitudes.
\end{enumerate}
Beyond these main hypotheses, we have formulated several sub-hypotheses that form the basis for our analysis plan (these sub-hypotheses also include several robustness checks and are available in \appref{app:AppendixHypotheses}).

@@

Given the more complex data structure of the extensive longitudinal data, we opted to test most of our hypotheses using multilevel regression models, where the measurement occasions (level 1) were nested within the participants (level 2). Given that more complex dynamic modeling procedures are not necessary to test these more basic hypotheses, we will focus on the contemporaneous effect of the data (e.g., need fulfillment during the intergroup interaction predicts outgroup attitudes following the interaction). Such an approach is also tolerant to missing data and uneven case numbers within participants. As is common with multilevel analyses, we use a hierarchical modeling approach and always report the final model in-text \citep[][; for the full modeling process see Supplementary Material B]{Bliese2013}.

@@

% Contact hypothesis in ESM data [inconsistent in aggregate but consistent in ML analysis]
When considering the results of the three studies jointly, we found mixed results for the basic intergroup contact hypothesis. More specifically, we observe that separating between-participant and within-participant effects is crucial. For between-subjects effects (using aggregated reports), the positive effect of more outgroup contact did not depend on the average interaction quality. These aggregate findings are surprising because most past cross-sectional studies have also considered an aggregate measurement of interaction quality (albeit not one from actual daily reports). The absence of this aggregate effect could underline the fact that cross-sectional retrospective data might be misleading because (1) it presents a mixture of within and between-subjects effects \citep[][]{Hamaker2020}, or (2) suffers from recall biases (e.g., where times with no outgroup interactions are undervalued by participants during the retrospective evaluations).\footnote{It should be noted that such inconsistencies with past research might in part be a data artifact (e.g., because most people reported substantially more measurements during which they did not have an outgroup interaction). However, the models we employed should in principle be resistant to such imbalances.} In support of this possibility, once we used multilevel modeling to separate effects into a between-subjects and within-subjects part, we found that when having an interaction with an outgroup member (vs. not having an outgroup interaction) is a strong predictor of outgroup attitudes across all three studies (even when taking other non-outgroup interactions into account; also see \fgrref[B]{fig:ContactHypothesis}). We thus find that the main benefits of intergroup contact remain visible across the three studies. However, the intensive longitudinal data of natural intergroup contacts also offers a novel opportunity to explore how people recall and cognitively aggregate outgroup interactions. Future studies may want to ask participants explicitly to evaluate the number and quality of outgroup interactions following a phase of intensive longitudinal data collection to disentangle the aggregation differences.

@@

\input{Tables/mdlContactGeneralLong}
\input{Tables/mdlTheorylong}

% Figures
\begin{figure}
  \caption{Path Diagrams Situational Needs Model across Studies}
  \label{fig:MainPaths}
  \begin{center}
    \input{PathModels.tex}
  \end{center}
  \caption*{Note: \\
  (A) Study 1; (B) Study 2; (C) Study 3.\\
  General: Path coefficients are standardized regression coefficients. Statistical significance markers are based on the unstandardized regression results (as presented in \tblref{tab:intergroupNeedsTblLong}).\\
  **** p < .0001, *** p < .001, ** p < .01, * p < .05}
\end{figure}

@@

\section{Hypotheses and Analysis Plan}
\label{app:AppendixHypotheses}

\newlength{\mdfmar}
\setlength{\mdfmar}{1.5em}
\mdfdefinestyle{mdfhypothesis}{
    innerleftmargin = +1.5\mdfmar, %
    innerrightmargin = +1.5\mdfmar, 
    innertopmargin = +\mdfmar, 
    innerbottommargin = +\mdfmar, 
    skipabove = 12pt
}
\newlength{\subhypskip}
\setlength{\subhypskip}{1.5em}
\newlength{\eqskip}
\setlength{\eqskip}{3.5em}

In this appendix, we present the expanded hypotheses and their associated analysis plan. Given the nested structure of much of our data we test many of our hypotheses using a multilevel approach, where $y_{ti}$ denotes the response at measurement occasion $t$ ($t = 1, ...,T_i$; level 1) for individual $i$ ($i = 1, ..., n$; level 2). All multilevel assumptions are tested as follows (e.g., for random slopes model with $j$ within-person predictors):

\begin{flalign}
  &\textrm{Level 1 Variance:}\ e_{ti} \sim \mathcal{N}(0,\sigma^2) \\
  &\textrm{Level 2 Variance:}\ \begin{bmatrix} u_{0i}\\ \vdots\\ u_{ji}\end{bmatrix} 
  \sim \mathcal{N}
  \begin{pmatrix}
    \begin{bmatrix} 
      0 \\ 
      \vdots \\
      0
    \end{bmatrix}, 
    \begin{bmatrix} 
      \tau_{00}^2 &  & \\ 
      \vdots & \ddots & \\
      \tau_{j0} & \ldots & \tau_{jj}^2
    \end{bmatrix}
  \end{pmatrix}
\end{flalign}

For our main aims we sequentially focus on four sets of models to test and validate our hypotheses:

\subsection{1. Contact hypothesis in extensive longitudinal data.} 
Within the individual studies, we begin by testing the most basic assumption of the intergroup contact hypothesis that outgroup attitudes should be more positive after outgroup interactions but not after non-outgroup interactions. For this we use multi-level regression analyses, predicting outgroup attitudes from outgroup interaction and non-outgroup interaction dummy variables. We also include the participant means as level two predictors to fully disentangle within- (level 1) and between-participant (level 2) effects of intergroup contact \citep[e.g.,][Section 4.6]{Snijders2012}.

\begin{mdframed}[style=mdfhypothesis]
    \begin{hyp}[H\ref{hyp:contactHyp}] \label{hyp:contactHyp}
        Based on the most general understanding of the contact hypothesis, positive intergroup contacts should be associated with more favorable outgroup attitudes across extensive longitudinal data.
    \end{hyp}

    \begin{subhyp}[H\ref{hyp:contactDummyML}] \label{hyp:contactDummyML}
    \addtolength{\leftskip}{\subhypskip}
    Outgroup attitudes should be more positive after an intergroup interaction compared to a non-outgroup interaction.
    \end{subhyp}

    \begin{fleqn}[\eqskip] 
    \begin{equation} \label{eq:ContactDummy}
      \begin{split}
        \textrm{Level 1:} &
          \begin{aligned}[t]
            \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}OutgroupInteraction_{ti} + \\
                               &\ \beta_{2i}NonOutgroupInteraction_{ti} + e_{ti}
          \end{aligned} \\
        \textrm{Level 2:} &
            \begin{aligned}[t]
                \ \beta_{0i} = &\ \gamma_{00} + \gamma_{01}MeanOutgroupInteraction_{i} + \\
                               &\ \gamma_{02}MeanNonOutgroupInteraction_{i} + u_{0i} \\
                  \beta_{1i} = &\ \gamma_{10} + u_{1i} \\
                  \beta_{2i} = &\ \gamma_{20} + u_{2i}
            \end{aligned}
      \end{split} 
    \end{equation}
    \end{fleqn}
\end{mdframed}

\hl{OPTION 1:} We then seek to investigate intergroup contacts and perceived interaction quality jointly. Notably, however, participants are only able to report their interaction quality perceptions if they had an interaction. This is considered to be structurally missing data and cannot meaningfully be imputed for modeling. To deal with this issue, we mimic the procedure commonly used within the cross-section literature. In particular, we conduct a person-level linear regression by aggregating the number of outgroup interactions participants had, their average interaction quality perceptions, as well as their average outgroup attitudes \citep[for possible alternative approaches see,][]{Enders2011}. To avoid an underpowered analysis, we use the aggregated data from all three studies.

\hl{OPTION 2:} We then seek to test the full contact hypothesis by investigating intergroup contacts and perceived interaction quality jointly. To do so, we conduct a person-level linear regression using person-level aggregated data from all three studies. In particular, we aggregate the number of outgroup interactions participants had, their average interaction quality perceptions, as well as their average outgroup attitudes. This approach has three main benefits: (1) Interaction quality ratings are only available if participants had an interaction, and the aggregation deals with this structural missingness. (2) Using the participant-level data from all three studies, we avoid potential power issues. (3) This analysis mimics the analyses conducted within the cross-section literature, where participants are asked to recall how many interactions they had over a one-month period, how positive these interactions were, and what their general attitudes towards the outgroup are.

\begin{mdframed}[style=mdfhypothesis]
    \begin{subhyp}[H\ref{hyp:contactCor}] \label{hyp:contactCor}
    \addtolength{\leftskip}{\subhypskip}
    Participants with more intergroup interactions should have more favorable outgroup attitudes.
    \end{subhyp}

      \begin{fleqn}[\eqskip] 
        \begin{equation} \label{eq:ContactCor}
            r_{\left(ContactFrequency_{i},\ AverageQuality_{i}\right)} > 0
        \end{equation}
      \end{fleqn}

    \begin{subhyp}[H\ref{hyp:contactQualLM}] \label{hyp:contactQualLM}
    \addtolength{\leftskip}{\subhypskip}
    Participants with more intergroup interactions should have more favorable outgroup attitudes depending on the average interaction quality.
    \end{subhyp}

      \begin{fleqn}[\eqskip] 
        \begin{equation} \label{eq:contactQualLM}
            \begin{split}
              AverageAttitude_{i} = &\ \beta_{0} + \beta_{1}ContactFrequency_{i} + \\
                                    &\ \beta_{2}AverageQuality_{i} +\\
                                    &\ \beta_{3}ContactFrequency_{i} * AverageQuality_{i}
            \end{split}
        \end{equation}
      \end{fleqn}

    \hl{More complete but in my opinion more confusing connotation:}

      \begin{fleqn}[\eqskip] 
        \begin{equation}
            \begin{split}
              \overline{Attitude}_{i} = &\ \beta_{0} + \beta_{1}\left(\sum\nolimits_{t=1}^{T}OutgroupInteraction\right)_{i} + \\
                                        &\ \beta_{2}\overline{InteractionQuality}_{i} +\\
                                        &\ \beta_{3}\left(\sum\nolimits_{t=1}^{T}OutgroupInteraction\right)_{i} * \overline{InteractionQuality}_{i}
            \end{split}
        \end{equation}
      \end{fleqn}

      We additionally control for the participant's study membership.
\end{mdframed}
Because this analysis uses the data from all three studies, the results of this analysis are presented in the `Robustness, Stability, and Embeddedness across Studies' section.

\subsection{2. Core need fulfillment during intergroup contact.}
The main proposal of this manuscript has been the assertion that the fulfillment of situation core needs during an interaction will be associated with more positive interaction quality perceptions and, ultimately, more positive outgroup attitudes. Thus, for the main set of analyses, we focus on the reported outgroup interactions only. For each study, we will use multi-level regressions to test the main three assertions of our proposal (mirroring the basic steps of a traditional mediation analysis).

\begin{mdframed}[style=mdfhypothesis]
    \begin{hyp}[H\ref{hyp:keyNeed}] \label{hyp:keyNeed}
    Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
    \end{hyp}

    \setcounter{subhyp}{0}
    \begin{subhyp}[H\ref{hyp:keyNeedPred}] \label{hyp:keyNeedPred}
    \addtolength{\leftskip}{\subhypskip}
    Core need fulfillment during outgroup interactions should predict more positive outgroup attitudes.
    \end{subhyp}

    \begin{fleqn}[\eqskip]
      \begin{equation} \label{eq:SlopesAttCore}
        \begin{split}
            \textrm{Level 1:} &\ Attitude_{ti} = \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + e_{ti}\\
            \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                              &\ \beta_{1i} = \gamma_{10} + u_{1i}
        \end{split}
      \end{equation}
    \end{fleqn}

    \begin{subhyp}[H\ref{hyp:keyNeedQual}] \label{hyp:keyNeedQual}
    \addtolength{\leftskip}{\subhypskip}
    Core need fulfillment during outgroup interactions should also predict higher perceived interaction quality.
    \end{subhyp}

    \begin{fleqn}[\eqskip]
      \begin{equation} \label{eq:SlopesQltCore}
        \begin{split}
            \textrm{Level 1:} &\ InteractionQuality_{ti} = \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + e_{ti}\\
            \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                              &\ \beta_{1i} = \gamma_{10} + u_{1i}
        \end{split}
      \end{equation}
    \end{fleqn}

    \begin{subhyp}[H\ref{hyp:keyNeedMediation}] \label{hyp:keyNeedMediation}
    \addtolength{\leftskip}{\subhypskip}
    The effect of core need fulfillment on outgroup attitudes should be reduced when considered together with perceived interaction quality.
    \end{subhyp}

    \begin{fleqn}[\eqskip]
      \begin{equation} \label{eq:SlopesAttCoreQual}
        \begin{split}
          \textrm{Level 1:} &
            \begin{aligned}[t]
              \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                                 &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
            \end{aligned} \\
          \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                            &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                            &\ \beta_{2i} = \gamma_{20} + u_{2i}
        \end{split} 
      \end{equation}
    \end{fleqn}
\end{mdframed}

\subsection{3. Allport's conditions in extensive longitudinal data.}
Within the third study, we formally measure all of Allport's optimal contact conditions. We use multi-level regression models to test whether the fulfillment of Allport's conditions in real-life data predicts more positive outgroup attitudes and higher perceived interaction quality, using the same approach as for the core need fulfillment above.

\begin{mdframed}[style=mdfhypothesis]
    \begin{hyp}[H\ref{hyp:Allport}] \label{hyp:Allport}
    Based on Allport's optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
    \end{hyp}

    \setcounter{subhyp}{0}
    \begin{subhyp}[H\ref{hyp:AttAllport}] \label{hyp:AttAllport}
    \addtolength{\leftskip}{\subhypskip}
    Higher fulfillment of Allport's conditions during outgroup interactions should predict more positive outgroup attitudes.
    \end{subhyp}

    \begin{fleqn}[\eqskip]
      \begin{equation} \label{eq:SlopesAttAllport}
        \begin{split}
            \textrm{Level 1:} &\ Attitude_{ti} = \beta_{0i} + \beta_{1i}Allport_{ti} + e_{ti}\\
            \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                              &\ \beta_{1i} = \gamma_{10} + u_{1i}
        \end{split}
      \end{equation}
    \end{fleqn}

    \begin{subhyp}[H\ref{hyp:QltAllport}] \label{hyp:QltAllport}
    \addtolength{\leftskip}{\subhypskip}
    Higher fulfillment of Allport's conditions during outgroup interactions should also predict higher perceived interaction quality.
    \end{subhyp}

    \begin{fleqn}[\eqskip]
      \begin{equation} \label{eq:SlopesQltAllport}
        \begin{split}
            \textrm{Level 1:} &\ InteractionQuality_{ti} = \beta_{0i} + \beta_{1i}Allport_{ti} + e_{ti}\\
            \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                              &\ \beta_{1i} = \gamma_{10} + u_{1i}
        \end{split}
      \end{equation}
    \end{fleqn}

    \begin{subhyp}[H\ref{hyp:AttAllportQual}] \label{hyp:AttAllportQual}
    \addtolength{\leftskip}{\subhypskip}
    The effect of higher fulfillment of Allport's conditions on outgroup attitudes should be reduced when considered together with perceived interaction quality.
    \end{subhyp}

    \begin{fleqn}[\eqskip]
      \begin{equation} \label{eq:SlopesAttAllportQual}
        \begin{split}
          \textrm{Level 1:} &
            \begin{aligned}[t]
              \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}Allport_{ti} + \\
                                 &\ \beta_{2i}InteractionQuality_{ti} + e_{ti}
            \end{aligned} \\
          \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                            &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                            &\ \beta_{2i} = \gamma_{20} + u_{2i}
        \end{split} 
      \end{equation}
    \end{fleqn}
\end{mdframed}

We then compare the effect of Allport's contact conditions with our core need fulfillment by comparing the model fit statistics of the two individual models and by adding both concepts to a joint multi-level regression model, to see whether the two approaches explain the same variance in outgroup attitudes.

\begin{mdframed}[style=mdfhypothesis]
    \begin{hyp}[H\ref{hyp:NeedAllport}] \label{hyp:NeedAllport}
    Based on our proposal, the fulfillment of the core situational need should predict outgroup attitudes at least as well as Allport's conditions.
    \end{hyp}

    \setcounter{subhyp}{0}
    \begin{subhyp}[H\ref{hyp:compModel}] \label{hyp:compModel}
    \addtolength{\leftskip}{\subhypskip}
    The need model (H\ref{hyp:keyNeedPred}) should predict more variance in outgroup attitudes than the model based on Allport's conditions (H\ref{hyp:AttAllport}).
    \end{subhyp}

    \begin{fleqn}[\eqskip] 
        \begin{equation}
            \begin{split}
                AIC_{KeyNeedModel} & < AIC_{AllportModel} \\
                BIC_{KeyNeedModel} & > BIC_{AllportModel}
            \end{split}
        \end{equation}
    \end{fleqn}

    \begin{subhyp}[H\ref{hyp:compTogether}] \label{hyp:compTogether}
    \addtolength{\leftskip}{\subhypskip}
    The  effect of key need fulfillment on outgroup attitudes should  persist even when taking other Allport's conditions into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for Allport's conditions.  
    \end{subhyp}

    \begin{fleqn}[\eqskip]
      \begin{equation} \label{eq:SlopesAttCoreAllport}
        \begin{split}
          \textrm{Level 1:} &
            \begin{aligned}[t]
              \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \\
                                 &\ \beta_{2i}Allport_{ti} + e_{ti}
            \end{aligned} \\
          \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                            &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                            &\ \beta_{2i} = \gamma_{20} + u_{2i}
        \end{split} 
      \end{equation}
    \end{fleqn}
\end{mdframed}

\subsection{4. Robustness, stability, and embeddedness across studies}
Within the final set of analyses, we look at the broader picture of our results and leverage the data from all participants to contextualize our results.

\paragraph{Robustness within studies}
To build further confidence in the effect of core need fulfillment during outgroup interactions, we conduct two additional robustness analyses for each study.

Firstly, to check for the role of alternate psychological needs, we add the fulfillment of self-determination theory needs (i.e., competence, autonomy, and relatedness) to the multi-level regression. We then also compare the model with models that predicts outgroup attitudes from self-determination theory need fulfillments or core need fulfillments only.

\begin{mdframed}[style=mdfhypothesis]
    \begin{hyp}[H\ref{hyp:keyNeedSDT}] \label{hyp:keyNeedSDT}
    The effect of key need fulfillment on outgroup attitudes should persist even when taking other fundamental psychological needs into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for autonomy, competence, and relatedness fulfillment during the interaction (cf., self-determination theory). 
    \end{hyp}

    \begin{fleqn}[\eqskip-\subhypskip]
      \begin{equation} \label{eq:SlopesAttCoreSdt}
        \begin{split}
          \textrm{Level 1:} &
            \begin{aligned}[t]
              \ Attitude_{ti} =  &\ \beta_{0i} + \beta_{1i}KeyNeedFulfill_{ti} + \beta_{2i}Autonomy_{ti} + \\
                                 &\ \beta_{3i}Competence_{ti} + \beta_{4i}Relatedness_{ti} + e_{ti}
            \end{aligned} \\
          \textrm{Level 2:} &\ \beta_{0i} = \gamma_{00} + u_{0i} \\
                            &\ \beta_{1i} = \gamma_{10} + u_{1i} \\
                            &\ \beta_{2i} = \gamma_{20} + u_{2i} \\
                            &\ \beta_{3i} = \gamma_{30} + u_{3i} \\
                            &\ \beta_{4i} = \gamma_{40} + u_{4i}
        \end{split} 
      \end{equation}
    \end{fleqn}
\end{mdframed}

To ensure that the core need fulfillment is outgroup contact specific, we return to the full sample of intensive longitudinal measurements within each study and test whether there is an interaction effect of outgroup contact (vs. no outgroup contact) and core need fulfillment. We expect that the effect of core need fulfillment is specific to outgroup interactions and not merely due to a more need-fulfilled life in general.

\begin{mdframed}[style=mdfhypothesis]
    \begin{hyp}[H\ref{hyp:keyNeedContactType}] \label{hyp:keyNeedContactType}
    \addtolength{\leftskip}{1em}
    The effect of key need fulfillment on outgroup attitudes should be specific to intergroup interactions and not be due to need fulfillment in general. Thus, the effect of key need fulfillment on outgroup attitudes should be stronger for intergroup interactions than for ingroup interactions. 
    \end{hyp}
